My wonderful paper

### 2024-12-03 _manuscript/_tex/index.tex
working on the linear regression application

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\section{Analysis plan}\label{sec-plan}

@@

@@

\textsubscript{Source:
\href{https://huizezhang-sherry.github.io/paper-analysis-plan/index.qmd.html}{Article
Notebook}}

@@

@@

\textsubscript{Source:
\href{https://huizezhang-sherry.github.io/paper-analysis-plan/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://huizezhang-sherry.github.io/paper-analysis-plan/index.qmd.html}{Article
Notebook}}

@@

Three examples are presented to illustrate how the concept of analysis
plan can be applied to data analysis. {[}toy example{]}.
Section~\ref{sec-linear-reg} illustrates how constructing the result
universe in a linear regression model of PM10 on mortality can help
understand the impact of sample size, model specification, and variable
correlation structure on data analysis. {[}example three{]}

@@

Consider a linear regression model to study the effect of PM10 on
mortality (provide context of using PM10 to study mortality). Analysts
may expect a significant (p-value \(\le\) 0.05) PM10 coefficient in the
linear model from the literature. This is the \emph{outcome
expectation}. There are multiple factors that can affect the outcome
expectation of linear regression, which here is called \emph{plan
expectation}, for example, 1) sample size, 2) model specification, and
3) correlation structure between variables. Adequate sample size is
required to achieve the desired power to detect the significance of PM10
on mortality. Temperature is often an important confounder to consider
in such study (add reference). From some domain knowledge, an analyst
may expect that the significance of PM10 coefficient can be attained by
adding temperature to the model. Analysts may also expect certain
correlation structure between PM10, temperature, and mortality, and the
distribution of each variable.

\begin{itemize}
\tightlist
\item
  add a paragraph to describe the simulation process
\end{itemize}

\textsubscript{Source:
\href{https://huizezhang-sherry.github.io/paper-analysis-plan/index.qmd.html}{Article
Notebook}}

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\section{xxxxx (not analysis plan)}\label{sec-plan}

@@

TODO: add one or two scenarios on why step count could go up or down: 1)
you take off your watch in the middle of the day, 2) have an workout

TODO: phrase it as Researcher measuring someone's step count rather than
measuring your own step count

@@

@@

top-down and bottom up (regression tree + logic tree): more naturally
useful summary of the say it is organized. (put down the plot)

@@

@@

A regression example is produced to illustrate the test selection
process for analyzing the effect of PM10 on mortality. The example
demonstrates how the test process can be used to select cutoff value in
the unit tests and how the test can be used to diagnose the unexpected
results.

@@

Consider a linear regression model to study the effect of PM10 on
mortality (provide context of using PM10 to study mortality). From the
literature, analysts may expect a significant (p-value \(\le\) 0.05)
PM10 coefficient in the linear model after considering the temperature
confounding {[}TODO add reference{]}. This expectation can be framed
into a unit test: the test fails, labelled as 1, if the p-value of the
PM10 coefficient is greater than 0.05. And we have devised three unit
tests to consider the 1) sample size, 2) strength of correlation between
mortality and PM10, and 3) strength of correlation between mortality and
temperature. {[}sample size matters because \ldots, correlation between
mortality and PM10 matters because\ldots{]}. However, it is not clear
how to set the cutoff value for these tests to determine a failure. Here
a list of tests are created in Table~\ref{tbl-tests}.

\begin{longtable}[]{@{}l@{}}
\caption{A list of unit tests created for the linear regression model of
mortality on PM10 and temperature. The tests are based on the sample
size, correlation between mortality and PM10, and correlation between
mortality and temperature. The cutoff values for each test are specified
to determine a failure.}\label{tbl-tests}\tabularnewline
\toprule\noalign{}
the test fails if \ldots{} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
the test fails if \ldots{} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
sample size less than or equal to 200 \\
sample size less than or equal to 400 \\
sample size less than or equal to 600 \\
sample size less than or equal to 800 \\
mortality-PM10 correlation less than -0.03 \\
mortality-PM10 correlation less than -0.04 \\
mortality-PM10 correlation less than -0.05 \\
mortality-PM10 correlation less than -0.06 \\
mortality-temperature correlation greater than -0.3 \\
mortality-temperature correlation greater than -0.35 \\
mortality-temperature correlation greater than -0.4 \\
mortality-temperature correlation greater than -0.45 \\
\end{longtable}

To generate replicate of the data, we first simulate the correlation
matrix for the three variables: mortality, PM10, and temperature with
the following vector:

\begin{itemize}
\tightlist
\item
  mortality and PM10: -0.03 to -0.1
\item
  mortality and temperature: -0.2 to -0.4
\item
  PM10 and temperature: 0.2 to 0.4
\end{itemize}

We then use a Gaussian copula to generate a multivariate normal
distribution based on the specified correlation matrix and sample size.
The multivariate normal distribution is transformed using the normal
CDF, followed by the inverse CDF of the assumed distributions of the
three variables. Various distributions are fitted to the variables,
including poisson and negative binomial for mortality; gamma,
log-normal, exponential, weibull, and normal for pm10 and temperature;
and beta for pm10 after rescaling the data to \([0-1]\). AIC is used to
determine the best distribution fit for each variable with qq plot
presented in Figure~\ref{fig-dist-fit} to evaluate the fit. The
following distributions and parameters, estimated from the original
data, are used to simulate the three variables:

\begin{itemize}
\tightlist
\item
  mortality, negative binomial distribution with size = 74 and mu = 183
\item
  PM10, beta distribution with shape1 = 4.21 and shape2 = 11.67, then
  multiply by 100
\item
  temperature, weibull distribution with shape = 3.8 and scale = 61
\end{itemize}

\phantomsection\label{cell-fig-dist-fit}
\begin{figure}[H]

\includegraphics{index_files/figure-pdf/fig-dist-fit-1.pdf}

\caption{\label{fig-dist-fit}QQ-plot of the distribution fit for
mortality, PM10, and temperature based on the fitted distribution from
the original data. The fitted distribution is compared to the observed
data to assess the distribution fit.}

\begin{figure}

\includegraphics{index_files/figure-pdf/unnamed-chunk-8-1.pdf}

\caption{\label{fig-linear-reg-tree}Logic regression model fitted to the
twelve unit tests and the outcome expectation (unexpect) as the response
variable. The model suggests the relationship:
(\(\text{cor}(\text{m}, \text{tmp}) > -0.35\) AND
\(\text{smpl\_size} \le 600\)) OR
\(\text{cor}(\text{m},  \text{PM10}) < - 0.05\)}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.0882}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1471}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1029}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1765}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1912}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1324}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1618}}@{}}

\caption{\label{tbl-linear-reg}Accuracy (precision and recall) and
parsimony (independence) metrics derived from the logic regression
model, along with harmonic and arithmetic means, for individual unit
tests (1: sample size, 2: mortality-PM10 correlation, 3:
mortality-temperature correlation), and the combined test rule 4:
(sample size AND mortality-temperature correlation) OR mortality-PM10
correlation.}

\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
tests
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
precision
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
recall
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
overlapping
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
independence
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
harmonic
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
arithmetic
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & 0.713 & 0.662 & 0 & 1 & 0.767 & 0.792 \\
2 & 0.906 & 0.748 & 0 & 1 & 0.872 & 0.885 \\
3 & 0.789 & 0.733 & 0 & 1 & 0.826 & 0.841 \\
4 & 0.908 & 0.794 & 0 & 1 & 0.893 & 0.901 \\

A total of 12 tests with four cutoff values each for the three types of
tests are used to fit the logic regression and calculate the precision
and recall score for predictability of an unexpected outcome (p value
\textless{} 0.05). Figure~\ref{fig-linear-reg-tree} shows the logic
regression tree from the fitted model and Table~\ref{tbl-linear-reg}
prints the numerical summary of three selected single test and their
combined test found by the logic regression. The logic regression model
is capable of picking up the best cutoff value for each type of tests
and yields the following three tests:

\begin{itemize}
\tightlist
\item
  sample size less than or equal to 600
\item
  mortality-temperature correlation greater than -0.35
\item
  mortality-PM10 correlation less than -0.05
\end{itemize}

The combined test rule, (mortality-tempearture correlation AND sample
size) OR mortality-PM10 correlation, has the highest precision, recall,
and independence, suggesting that it is the most effective test for
diagnosing unexpected results in the linear regression model given our
assumption on the data. The tree structure suggests obtaining the
mortality-PM10 correlation is as important as obtaining a combined
relationship between mortality-temperature correlation and sample size
in diagnosing unexpected results in the linear regression model. And
based on our assumption of the variable distribution from the data
simulation, we can obtain a precision of 0.908 and a recall of 0.794 for
the unexpected outcome. These provides check points for the analysts to
diagnose the unexpected results and to improve the model.
